### 2025-01-09 doc/paper.tex
Work on Introduction

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

\usepackage{tikz}
% Corporate Design of the University of Tübingen
% Primary Colors
\definecolor{TUred}{RGB}{165,30,55}
\definecolor{TUgold}{RGB}{180,160,105}
\definecolor{TUdark}{RGB}{50,65,75}
\definecolor{TUgray}{RGB}{175,179,183}

% Secondary Colors
\definecolor{TUdarkblue}{RGB}{65,90,140}
\definecolor{TUblue}{RGB}{0,105,170}
\definecolor{TUlightblue}{RGB}{80,170,200}
\definecolor{TUlightgreen}{RGB}{130,185,160}
\definecolor{TUgreen}{RGB}{125,165,75}
\definecolor{TUdarkgreen}{RGB}{50,110,30}
\definecolor{TUocre}{RGB}{200,80,60}
\definecolor{TUviolet}{RGB}{175,110,150}
\definecolor{TUmauve}{RGB}{180,160,150}
\definecolor{TUbeige}{RGB}{215,180,105}
\definecolor{TUorange}{RGB}{210,150,0}
\definecolor{TUbrown}{RGB}{145,105,70}

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2023} with \usepackage[nohyperref]{icml2023} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

\usepackage[accepted]{icml2023}

% For theorems and such
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}

% if you use cleveref..
\usepackage[capitalize,noabbrev]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
%\usepackage[disable,textsize=tiny]{todonotes}
\usepackage[textsize=tiny]{todonotes}
\usepackage{float}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Comparing different active learning query methods}

\begin{document}

\twocolumn[
\icmltitle{TODO: Funky active learning pun}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2023
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
	\icmlauthor{Lukas Weber}{first}
\end{icmlauthorlist}

\icmlaffiliation{first}{MSc Computer Science, Tübingen, GER}
\icmlcorrespondingauthor{Lukas Weber}{lukas2.weber@student.uni-tuebingen.de}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Machine Learning, Active Learning}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
	A lot of unlabelled data generally available but annotation is costly. A goal in fields where labelled data is not available  is to maximize performance with minimum amount of labeled data instances. \\
	Active learning selects instances for labelling based on a selection strategy. 
	\begin{itemize}
		\item Objective: Compare active learning query methods using modAL.
		\item Methods: Experiments with Logistic Regression (MNIST) and a CNN (CIFAR).
		\item Experiments: Vary initial dataset sizes (small, moderate, large).
		\item Key findings: Highlight main conclusions about performance in different scenarios.
	\end{itemize}
\end{abstract}

\section{Introduction}\label{sec:intro}
Complex tasks in supervised learning usually require large annotated datasets. However, the cost and time needed to label samples can be high, especially when expert knowledge in the according field is required.  \\
Active learning aims to reduce the overall labeling cost by selecting the most informative samples for annotation by an oracle and thus enable using machine learning for tasks where annotated data is sparse. \\
The effectiveness of active learning relies heavily on the choice of query strategy, as it determines which samples are selected for annotation and thus directly influences model performance. In pool-based active learning, a model is trained on a small labeled dataset and used to evaluate a large pool of unlabeled data. Informative samples are selected based on a query strategy, labeled by an oracle, and added to the training set. The model is retrained, and the process is repeated until a stopping criterion is met. \\ 
In stream-based active learning, unlabeled data arrives sequentially. Thus, the model decides whether to query the label of an incoming instance using an evaluation metric like the uncertainty of the prediction and checking whether a threshold is reached. If not, the instance is discarded. \\
This report focuses on comparing several pool-based as well as stream-based query strategies. The modular Active Learning framework (modAL)~\cite{danka_modalmodularactivelearning} is used in the following experiments as it simplified the implementation of active learning pipelines, contains multiple already implemented querying strategies and is compatible with models from the Scikit-learn library~\cite{scikit-learn}. \\
Several previous works explore different query methods for different tasks. \\
\cite{schröder_revisitinguncertaintybasedquerystrategies} compares various uncertainty based query strategies in the context of fine-tuning transformer models for text classification, while \cite{zhan_comparativesurveydeepactive} compares several query methods on MNIST and CIFAR. \\
According to \cite{ueno_benchmarkingofquerystrategies} most papers concentrate on two image-based datasets: MNIST~\cite{mnist} and CIFAR~\cite{cifar}, while their paper evaluates strategies on six different datasets, including medical and visual inspection images. They also verify the effectiveness of query strategies independently of the training process, underfitting and randomness due to early-stage training in an experiment where fully trained models are used to select the most informative samples according to various query strategies and to construct a labeled dataset. \\ 
\cite{werner_comparableactivelearning} agrees with their statement that experimental settings are not standardized across papers which makes comparisons of different existing and new query strategies difficult. They propose a benchmark framework evaluating active learning strategies on tabular, image, and text datasets using robust evaluation protocols to mitigate variance and ensure comparability. \\
While \cite{ueno_benchmarkingofquerystrategies} evaluates strategies across several image-based datasets from a wide range of image-domains like the medical and visual inspection images, \cite{werner_comparableactivelearning} focuses on a wide range of data-domains including images, text and vector-based datasets. \\
This report focuses on comparing strategies in the image domain. It compares query strategies for a simple logistic regression model and a more complex CNN to study the effect of model complexity on the optimal query strategy. \textcolor{red}{CITATION were able to show that [QUERY METHODS] work well on simple models but do not generalize well on more complex models. To guarantee real-world applicability, knowledge about this needs to be gained as model complexities vary across different tasks and datasets}. Furthermore, the impact of different sizes of the initial annotated dataset is additionally investigated which is something that few studies analyze in combination. The effect of the combination between model complexity and initial annotated dataset size is a research gap in previously mentioned papers. \\
By examining different starting sizes, this work simulates various real-world scenarios: (1) starting with no labeled data, (2) having a small pool of labeled data, and (3) working with a well-trained model to identify data points for fine-tuning.
% FROM HERE THIS IS NOT GOING TO BE PART OF THE FINAL PART
Previous papers do not: Is everything here now mentioned previously? if yes, then it can be deleted.
\begin{itemize}
	\item \cite{schröder_revisitinguncertaintybasedquerystrategies}: Does not compare across different sizes of the pool, and for simple models.
	\item \cite{comparativesurveydeepactive}: Does not compare impact of different initial train set size and strategies on shallow classifiers.

	\item According to \cite{ueno_benchmarkingofquerystrategies}: \textcolor{red}{check if true:} Does not compare different sizes of the initial starting sets as well as how simple and more complex perform with the query methods

	\item \cite{werner_comparableactivelearning}: They focus on a wide range of domains, this paper only compares methods in the image domain. I compare strategies for a simple logistic regression model and a complex CNN across different starting sizes of the annotated dataset. \\
\end{itemize}
% TO HERE
The contributions of this paper are as follows:
\begin{itemize}
	\item \textbf{Systematic evaluation of active learning query strategies} on MNIST with a logistic regression model and CIFAR-10 using a CNN to analyze their behavior across datasets and model complexities.
	\item \textbf{Investigation of model complexity's influence} on the effectiveness of query strategies to provide knowledge into how shallow and deep models respond differently to active learning methods.
	\item \textbf{Analysis of the initial labeled dataset size's impact} on the performance of query strategies to simulate real-world scenarios such as starting completely from scratch, having a small labeled dataset or fine-tuning a model.
\end{itemize}

\textcolor{red}{Should I even give a structure of the paper?}

\section{Methods}\label{sec:methods}
\label{data_and_methods}
\subsection{Datasets}
\begin{itemize}
	\item MNIST and CIFAR describe
	\item processing steps \\
	MNIST: Vectorized, flattened (describe input size), normalized
\end{itemize}

\subsection{Models}
\begin{itemize}
	\item Logistic Regression: Details on implementation and hyperparameters, how were they chosen?
	\item same for CNN
\end{itemize}

\subsection{Active Learning}
\begin{itemize}
	\item How are the splits (initial pool, initial training data) generated
	\item numbmer of iterations for active learning
	\item Used active learning query strategies
\end{itemize}

Metrics for evaluation (Accuracy, train error, confusion matrix, ...)

\section{Experiments}\label{sec:experiments}
Setup: details for reproducibility (random number generators)?
Random seed = 42
\cite{comparableactivelearning} mention reproducibility problems in models even when setting random number generators before each run. By training models multiple times in succession and checking whether the same configurations lead to different outcomes, it is shown that in this scenario, reproducibility is guaranteed. \textcolor{red}{TODO: Schnell ein skript aufsetzen und dann finale accuracy und so weiter vergleichen.}
\\
Did I perform multiple runs?
\\

Scenarios
\begin{itemize}
	\item describe experiments and their settings and purposes
\end{itemize}

\section{Results}\label{sec:results}
Quantitative results: Tables or plots comparing the query strategies across scenarios. Also maybe key perforamnce differences (accuracy vs number of labeled examples)

\section{Discussion}\label{sec:discussion}
Qualitative analysis: Why certain strategies perform better or worse in specific settings. Trends between simple models and complex models
\\
\\
Impact of initial dataset size
\\
\\
Most methods require knowledge about the balance of the dataset and might be disadvantageous to use in case of a disbalanced dataset. In real world scenarios, this is not known. \\
Only image-based datasets were tested. Not vector- or text-based.

\section{Conclusion}\label{sec:conclusion}
Summary of findings and insights (maybe that a strategy works best with small datasets while another one is robust on bigger ones)
\\
Implications: practical recommendations?
\\
Future work? (More diverse datasets, exploring additional strategies, applying in real-world tasks)

\section{Cool ideas}
\begin{itemize}
	\item Show the typical framework of active learning in a figure in the introduction
\end{itemize}

\section{Good formulations}
From the active learning survey: Entropy is an information-theoretic
measure that represents the amount of information needed to “encode” a distri-
bution. As such, it is often thought of as a measure of uncertainty or impurity in machine learning.

\bibliography{bibliography}
\bibliographystyle{icml2023}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2021 and by Csaba Szepesvari, Gang Niu and Sivan Sabato in 2022.
% Modified again in 2023 by Sivan Sabato and Jonathan Scarlett.
% Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.